### 2023-06-13 state-resilience.tex
up

Resilience is a strong property since it implies that from every element of $\Bad$ there must exist a path to $\Safe$. However, when one considers a system with an initial state $s_0$, it could be sufficient to ask that only from $\Bad \cap \post^*(s_0)$, there must exist a path to $\Safe$. 
%
%			However, this condition seems more difficult since one also must decide whether a state $s \in \Bad$ is reachable from $s_0$.
%
%\problemx{(I,J)-$k$-resilience problem for WSTS}
%{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis), a set $J$.}
%{$\forall s' \in J ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
%
%\problemx{general $k$-resilience problem for WSTS}
%{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis).}
%{$\forall s'  ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}

%		Again, we first consider the case where the recovery time is bounded by a $k \in \N$.

The three previous problems become:

@@

Since these problems are undecidable for general infinite-state transition systems, we restrict our study to WSTS.

As in the Section~3, we study decidability results for different pairs of sets $\Safe$ and $\Bad$ downward-closed and upward-closed. We will consider here the cases where 
$\Safe = \uparrow \Safe$ and $\Bad = \downarrow \Bad$ 
and where 
$\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.

@@

Reset VASS extend the basic VASS model with special “reset
transitions” that resets (set to $0$) some coordinates in the vector. Let us recall their definition here.\\

% More Formally {\em reset ... }

\begin{definition}
A {\em Reset VASS} in dimension $d$ (Reset $d$-VASS for short) is a finite 
% $\mathds{Z}^d$
labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em control-states} of $V$, and where 
%$T \subseteq Q \times \mathds{Z}^d \times Q$
$T \subseteq Q \times Op \times Q$
 will be referred to as the {\em control-transitions} of $V$,
where $Op = \{ add(\textbf{z}) \mid \textbf{z} \in \mathds{Z}^d\} \cup 
		\{ reset(i) \mid i \in \{1,\ldots,d\} \}$.
\end{definition}

Again $Q \times \N^d$
 denotes the set of configurations of $V$.
For every configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control-transition $t$ we write
$p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ when

\begin{itemize}

\item  $t = (q,add(\textbf{z}),q') \in T$
% then for all $\textbf{u} \in \N^d$ such that  
% $\textbf{u}+\textbf{z} \geq 0$
% $q(\textbf{u}) \xrightarrow{\textbf{z}} q'(\textbf{u}+\textbf{z})$,
and $\textbf{u}+\textbf{z} = \textbf{v} \geq 0$,

\item $t = (q,reset(\gamma),q') \in T$ 
% then for all $\textbf{u} \in \N^d$ 
% $q(\textbf{u}) \xrightarrow{z} q'(\textbf{u}')$,  where 
and
$\textbf{v}[\gamma] = 0$, and $\textbf{v}[\gamma'] = \textbf{u}[\gamma']$ for all $\gamma' \in \{1,\ldots, d\} \setminus \gamma$.
\end{itemize}

\iffalse
\problemx{$d$-VASS zero-reachability}
{VASS $V=(Q,T)$, $p(\textbf{u}) \in Q \times \N^d$}
{$\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$? \\} 
\fi

It is well known that Reset $d$-VASS are WSTS [ref ? relire WSTS everywhere]. 
Moreover, we are interested in the 
%following
 undecidable \cite{araki1976PN} decision problem%:
\iffalse
\alain{ce qui suit est inutile}

@@

\iffalse
\alain{definir downward reachability problem.....
downward-closed problem given a state $s$ of a WSTS
% in the regarded class 
with strong compatibility 
and a decidable downward-closed set $D$, it can be decided whether $\exists s' \in D ~ s \to^* s'$. }
\fi

>>>>

Resilience is a strong property since it implies that from every element of $\Bad$ there must exist a path to $\Safe$. However, when one considers a system with an initial state $s_0$, it could be sufficient to ask that only from $\Bad \cap \post^*(s_0)$, there must exist a path to $\Safe$. 
%
%			However, this condition seems more difficult since one also must decide whether a state $s \in \Bad$ is reachable from $s_0$.
%
%\problemx{(I,J)-$k$-resilience problem for WSTS}
%{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis), a set $J$.}
%{$\forall s' \in J ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
%
%\problemx{general $k$-resilience problem for WSTS}
%{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis).}
%{$\forall s'  ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
%		Again, we first consider the case where the recovery time is bounded by a $k \in \N$.
The three previous problems become:

@@

Since these problems are undecidable for general infinite-state transition systems, we restrict our study to WSTS.
As in the Section~3, we study decidability results for different pairs of sets $\Safe$ and $\Bad$ downward-closed and upward-closed. We will consider here the cases where 
$\Safe = \uparrow \Safe$ and $\Bad = \downarrow \Bad$ 
and where 
$\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.

@@

Reset VASS extend the basic VASS model with special “reset
transitions” that resets (set to $0$) some coordinates in the vector. Let us recall their definition here.
\begin{definition}
A {\em Reset VASS} in dimension $d$ (Reset $d$-VASS for short) is a finite 
% $\mathds{Z}^d$
labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em control-states} of $V$, and where 
%$T \subseteq Q \times \mathds{Z}^d \times Q$
$T \subseteq Q \times Op \times Q$
 will be referred to as the {\em control-transitions} of $V$,
where $Op = \{ add(\textbf{z}) \mid \textbf{z} \in \mathds{Z}^d\} \cup 
		\{ reset(i) \mid i \in \{1,\ldots,d\} \}$.
\end{definition}

Again $Q \times \N^d$
 denotes the set of configurations of $V$.
For every configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control-transition $t$ we write
$p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ when 
\begin{samepage}\begin{itemize}
\item  $t = (q,add(\textbf{z}),q') \in T$
% then for all $\textbf{u} \in \N^d$ such that  
% $\textbf{u}+\textbf{z} \geq 0$
% $q(\textbf{u}) \xrightarrow{\textbf{z}} q'(\textbf{u}+\textbf{z})$,
and $\textbf{u}+\textbf{z} = \textbf{v} \geq 0$,
\item $t = (q,reset(\gamma),q') \in T$ 
% then for all $\textbf{u} \in \N^d$ 
% $q(\textbf{u}) \xrightarrow{z} q'(\textbf{u}')$,  where 
and
$\textbf{v}[\gamma] = 0$, and $\textbf{v}[\gamma'] = \textbf{u}[\gamma']$ for all $\gamma' \in \{1,\ldots, d\} \setminus \gamma$.
\end{itemize} \end{samepage}

It is well known that Reset $d$-VASS are WSTS \cite{dufourd1998reset}. 
Moreover, we are interested in the 
%following
 undecidable \cite{araki1976PN} decision problem%:
\iffalse
\alain{ce qui suit est inutile}

@@
